Safe Memory Regions for Big Data Processing

### 2016-05-30 paper/type-inference.tex
Added high-level outline of type-inference

The elaboration consists of the following steps:
\begin{enumerate}
 \item The first step consists of the following two mutually dependent steps:
   \begin{enumerate}
     \item for every class determine a set of region variable parameters ($\rho$)
       it should be parametrized over, and
     \item for every field of the class, determine its region parametrized type by
       generating as many fresh region variables as required by its class (type).
   \end{enumerate}
 \item  For every method and function type, determine its corresponding region
   type template by introducing fresh region variable parameters as required by
   the type of each parameter and return value (in addition to the allocation region
   parameter).
 \item Elaborate every \C{new} expression, method call or function application by
   introducing fresh region variable parameters (as required by the region type
   template of the called method or function).
 \item Introduce for each class and method, a predicate variable ($\varphi$) to denote
   an undetermined set of constraints over the region variable parameters of that
   class/method.
 \item We analyze each method, to generate a set of constraints that capture
   well-formedness requirements and subtyping relationships between type templates
   that must hold (as per the static semantics in Fig.~\ref{fb-staticsem}).
 \item We solve the generated set of constraints using our fixpoint constraint
   solving algorithm \csolvestar, which reduces the constraint solving problem to
   an abduction problem in a Herbrand constraint system, and then relies on \csolve,
   our abduction solver for that domain. 
 \item We finally simplify the elaborated program by coalescing region variables
   determined to be the same in the constraint solution.
\end{enumerate}

>>>>

Fig.~\ref{fig:type-inference-algo} presents the high-level outline of the type
inference algorithm.
The algorithm consists of the following steps:
\begin{enumerate}
 \item The first step elaborates the input program by introducing \emph{(formal and actual) region parameters},
and \emph{region variables} (representing yet undetermined \emph{actual region parameters})
as follows:
\begin{enumerate}
\item For every class determine a set of \emph{region parameters} ($\rho$)
       it should be parametrized over.
\item For every field of the class, determine its region parametrized type by
       generating as many fresh \emph{actual region parameters} as required by its class (type),
       which in turn become formal region parameters of the containing class.
\item  For every method and function type, determine its corresponding region
   type template by introducing fresh region parameters as required by
   the type of each parameter and return value (in addition to the allocation region
   parameter).
 \item Elaborate every \C{new} expression, method call or function application by
   introducing \emph{fresh region variables} (as required by the region type
   template of the called method or function).
\item Introduce for each class and method, a \emph{predicate variable} ($\varphi$) to denote
   an undetermined set of constraints over the region parameters of that
   class/method.
\end{enumerate}
 \item In the second step, we analyze the program to generate a set of constraints that capture
   well-formedness requirements and subtyping relationships between type templates
   that must hold (as per the static semantics in Fig.~\ref{fb-staticsem}).
 \item We solve the generated set of constraints using our fixpoint constraint
   solving algorithm \csolvestar, which reduces the constraint solving problem to
   an abduction problem in a Herbrand constraint system, and then relies on \csolve,
   our abduction solver for that domain. 
 \item We finally simplify the elaborated program by substituting the solution determined
   (for the region variables and predicate variables).
\end{enumerate}

\begin{figure}
\begin{numcodeml}
Infer ($p$) =
  let $q$ = IntroduceVariables($p$) in
  let $(r,C)$ = GenerateConstraints($q$) in
  match (SolveConstraints($C$) with
  | None $\longrightarrow$ None
  | Some ($s$) $\longrightarrow$ Some (ApplySolution $(s,r)$)
\end{numcodeml}

\caption{The type inference algorithm}
\label{fig:type-inference-algo}
\end{figure}

\begin{theorem}
\emph{(\textbf{Soundness})}
For any $p \in \absof{\FB}$, if Infer($p$) returns Some($t$), then
(1) $\absof{t} = p$, and
(2) $t$ is well-typed.
\end{theorem}
